Numerical trajectory patterns of floating macroalgae

### 2020-09-19 thesis_chapter_3.Rmd
updated chapter 3 and chpater 1. general comits and temporary track data added

@@

The characteristics of the underlying numerical flow model will be explained here...
## Particle tracking model

@@

$$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model, $X_b(t)$ is a change in position due to 'behavior' of the particle. This can range from swimming in fish to sinking or beaching. IN order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior. This study simulated two scenarios; one where the plant is fully submerged and only exposed to seawater drag and the other where parts of the plants are partially exposed resulting in a combination of seawater and air form drag. Unlike previous studies, drag based on shape based coefficients was used in the calculations for determining the reduction in velocity as a result of drag. A momentum energy equation was used to calculate the drag forces for the relevant simulation. This approach has been employed in modeling iceberg drift trajectory [@andersson2017; @lichey2001; @eik2009] and was adapted to suit this particular study. The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was, $$m({d_u/d_t})=F_c + F_a + F_f$$
where $m$ is kelp mass, $d_u$ the virtual kelp particle velocity, $F_c$ the hydrodynamic drag, $F_a$ is the wind drag, and $F_f$ the surface-current flow. The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations. Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculated before input into the model.

@@

@@

>>>>

## Study site

```{r map}

## Load software:
library(ggplot2)
library(rgdal)
library(tidyverse)
library(viridis)
library(gridExtra)
source("map/scaleBarFunc.R")

load("map/south_africa_coast_hi_res.Rdata")

map_pen <- ggplot(data = south_africa_coast_hi_res) +
geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
            size = 1.0, colour = "black", fill = NA) +
  #geom_point(data = study_sites, aes(x = lon, y = lat), color = "red", nudge_x = 0, nudge_y = 0.1) +
  #geom_label(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
  geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.10, xmax=18.50, ymin=-33.70, ymax=-34.50), inherit.aes = FALSE, fill = "blue", alpha = 0.01) +
  geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.40, xmax=18.90, ymin=-34.00, ymax=-34.50), inherit.aes = FALSE, fill = "red", alpha = 0.01) +
  #annotate("text", label = "Warm\nTemperate", x = 18.7, y = -34.5, size = 4.0, angle = 0, colour = "black") +
  #annotate("text", label = "Cool\nTemperate", x = 18.0, y = -34.0, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "FALSE\nBAY", x =18.65, y = -34.25, size = 4.0, angle = 0, colour = "black") +
  scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
  arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
  #theme(axis.ticks = element_blank()) + # Remove tick marks
  xlab("Latitude") +
  ylab("Longitude")

## Coastline of Southern Africa
load("map/south_africa_coast_hi_res.Rdata")

release_sites <- read.csv("metadata/release_sites.csv", sep = ";")

## Make 2nd map of just the peninsula
release_map <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  geom_point(data = release_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
  geom_label(data = release_sites, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -33.600)) +
  theme(axis.ticks = element_blank(), # Remove tick marks
        #axis.text = element_blank(), # Remove lat/ lon numbers
        axis.title = element_blank()) # Remove lat/ lon labels
#ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")

```

```{r}

library(ggpubr)

ggarrange(map_pen, release_map, labels = c("AUTO"), align = c("hv"), ncol = 2, nrow = 1)

```

@@

The characteristics of the underlying numerical flow model will be explained here...

## Particle tracking model

@@

$$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model, $X_b(t)$ is a change in position due to 'behavior' of the particle. This can range from swimming in fish to sinking or beaching. IN order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.

This study simulated two scenarios; one where the plant is fully submerged and only exposed to seawater drag and the other where parts of the plants are partially exposed resulting in a combination of seawater and air form drag. Unlike previous studies, drag based on shape based coefficients was used in the calculations for determining the reduction in velocity as a result of drag. A momentum energy equation was used to calculate the drag forces for the relevant simulation. This approach has been employed in modeling iceberg drift trajectory [@andersson2017; @lichey2001; @eik2009] and was adapted to suit this particular study. The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was, $$m({d_u/d_t})=F_c + F_a + F_f$$
where $m$ is kelp mass, $d_u$ the virtual kelp particle velocity, $F_c$ the hydrodynamic drag, $F_a$ is the wind drag, and $F_f$ the surface-current flow. The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations. Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculated before input into the model.

@@

## Windage

In order to calculate the direct effect of wind-drag on the kelp the wind velcocity is needed, which can be estimated by the surface-current velocity.

The total wind drag effect was calculated for the hypothetical cross-sectional area of the kelp particle which was exposed above the water surface, which in this study, was estimated as 10% of the total cross-sectional area.

@@

```{r}

# Read in data and data wrangling

library(tidync)
library(tidyverse)

p_particle <- tidync("data/trajectory_data/PassiveFloat.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))

p_particle_mean <- p_particle %>% 
   group_by(t) %>% 
   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)

p_particle$type <- "passive" # add extra column 'type' to dataframe and assigned 'passive' label

k_particle <- tidync("data/trajectory_data/KelpFloat.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))

k_particle_mean <- k_particle %>% 
   group_by(t) %>% 
   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)

k_particle$type <- "kelp" # add extra column 'type' to dataframe and assigned 'kelp' label

total_data <- rbind(p_particle, k_particle)

```

```{r}

plot_A <- ggplot(data = p_particle, aes(x = lon, y = lat)) + 
  geom_point() + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  theme_bw()

plot_B <- ggplot(data = k_particle, aes(x = lon, y = lat)) + 
  geom_point() + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  theme_bw()

```

```{r}

library(ggpubr)
ggarrange(plot_A,plot_B, labels = c("A","B"))

```
